---
title: "Simulations for postdoc secretary problem"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
Code for simulating one experient, where we reject $k_0$ out of $n$ and then accept the first, who is second-best:
```{r}
simulate <- function(n, k0){
  lst <- sample(1:n, n)
  first <- lst[1]
  second <- 0
  for(i in 2:k0){
    if (lst[i] < first){
      second <- first
      first <- lst[i]
    }else if(lst[i] < second){
      second <- lst[i]
    }
  }
  for(i in (k0+1):n){
    if (lst[i] < first){
      second <- first
      first <- lst[i]
    }else if(lst[i] < second){
      if (lst[i] == 2){
        return(1);
      }else{
        return(0);
      }
    }
  }
  return(0);
}

simulate_N_times<- function(N, n, k0){
  result<- 0
  for (i in 1:N){
    result <- result + simulate(n, k0)
  }
  return (result/N)
}
```
Run simulation $N$ times for different values of $n$(ranging from $50$ to $1000$ with the step $4$):
```{r}
N <- 700
n_vect <- seq(50, 300, by=2)
max_success <- c()
argmax_success <- c()
j <- 1
for (n in n_vect){
  success <- c()
  for (k0 in 1:(n - 1)){
    success[k0] <- 0
    for (i in 1:N){
      success[k0] <- success[k0] + simulate(n, k0)
    }
  }
  max_success[j] <- max(success) / N
  argmax_success[j] <- which.max(success)
  j <- j + 1
}
```
Plot the dependence of $n$ and the maximum probability for that $n$ and also at what value of $k_0$ maniizes the probability:
```{r}
plot(n_vect, max_success, type='l', xlab="n", ylab="Probability")
```


```{r}
plot(n_vect, argmax_success, xlab="k0", ylab="Probability")
lines(n_vect, n_vect / 2)
```
As we can see, maximum probability of success following optimal strategy is near $0.25$ (actually it is a little bigger on the graph, since our number of experiment is a finite number) and is achived by the value of $k_0$ that is close to $\frac{n}{2}$ as was suggested by theory.  


Code for calculating success probability using derived formula:
```{r}
calculate <- function(n, k0){
  return(k0 * (n - k0)/(n * (n - 1)))
}
```
Calculate the difference between result of simulations and results from derived formula:
```{r}
n <- 100
difference <- c()
formula <- c()
simulation <- c()
for (k0 in 1:(n - 1)){
  formula[k0] <- calculate(n, k0)
  simulation[k0] <- simulate_N_times(1000, n, k0)
  difference[k0] <- abs(formula[k0] - simulation[k0])
}
cat("Max difference for success relative frequencies and success probabilities is: ", max(difference), end="\n")
plot(1:(n-1), formula, type='l')
lines(1:(n-1), simulation)
```
Code for determining the value of $k_0$ that minimizes $E(R_n)$:
```{r}
simulate_rank <- function(n, k0){
  rank <- 0
  lst <- sample(1:n, n)
  first <- lst[1]
  second <- 0
  for(i in 2:k0){
    if (lst[i] < first){
      second <- first
      first <- lst[i]
    }else if(lst[i] < second){
      second <- lst[i]
    }
  }
  for(i in (k0+1):n){
    if (lst[i] < first){
      second <- first
      first <- lst[i]
    }else if(lst[i] < second){
      return (lst[i])
    }
  }
  return(lst[length(lst)]);
}

M <- 1000
N_values <- seq(50, 300, by=10)
min_n = c()
for (N in N_values){
  results <- c()
  for (k0 in 1:(N-1)){
    rank = 0
    for (m in 0:M){
      rank <- rank + simulate_rank(N, k0)
    }
    results[k0] = rank/M
  }
  min_n <- c(min_n, which.min(results))
}

plot(N_values, min_n, xlab = "N - length of sequence", ylab="n - stop value for min E(R)")

```
Code for simulating secretary problem, when we can choose either first or second and it will count as a success:
```{r}
simulate <- function(n){
  m1 = 0.347 * n
  m2 <- 2/3 *n
  rank <- 0
  lst <- sample(1:n, n)
  first <- lst[1]
  second <- 0
  for(i in 2:m1){
    if (lst[i] < first){
      second <- first
      first <- lst[i]
    }else if(lst[i] < second){
      second <- lst[i]
    }
  }
  for(i in (m1+1):m2){
    if (lst[i] < first){
      if ((lst[i]==1) | (lst[i]==2)){
        return (1)
      } else {
        return (0)
      }
    } else if (lst[i]< second){
      second <- lst[i]
    }
  }
    for(i in (m2+1):n){
    if (lst[i] < second){
      if ((lst[i]==1) | (lst[i]==2)){
        return (1)
      } else {
        return (0)
      }
    }
  }
  return (0);
}

N_values <- c(100, 500, 1000, 5000)
M <- 1000

success = 0
for (n in N_values){
  success = 0
  for (m in 1:M){
  success = success + simulate(n)
}
cat("Probability of success for n =", n, "is", success/M, '\n')
}

```

